\documentclass[a4paper,12pt]{article}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}
\title{Worksheet 6}
\author{Conor Houghton, copying from Martha Lewis}
\date{}

\begin{document}

\maketitle

\section{Introduction}

This worksheet is a little different from the previous ones: there is
no programming! Instead, the idea is for you to read about a dataset
and create a datasheet for it, using the ideas from Gebru et
al. \cite{gebru2021datasheets}.

We will be using the datasheet questions from Gebru et
al. \cite{gebru2021datasheets}, and applying them to the ImageNet
dataset \cite{deng2009imagenet} to create a datasheet for the ImageNet
dataset. If you cannot find the answers to all the questions, don’t
worry! The idea is for you to start thinking about the ideas and also
to see the extent to which the information needed for the datasheet is
publicly available.

\section{Task}

First of all, open up the paper \textit{Datasheets for Datasets}. The
URL is here:\\
\\
\url{https://arxiv.org/abs/1803.09010}.\\
\\
The set of questions that the authors propose should be used is in
Sections 3.1-3.7. In each section, pick two or three of the questions
and try to answer them in relation to the ImageNet dataset (paper
here:\\ \\ \url{http://www.image-net.org/papers/imagenet_cvpr09.pdf}\\ \\ and
website here: \url{http://www.image-net.org/}).

At the end of the \textit{Datasheets for Datasets} paper, there are two example datasheets that you can look at to get ideas. Feel free to work together if it is helpful!

\begin{itemize}
    \item How easy is it to answer the questions?
    \item Is one category of questions easier to answer than another?
    \item How easy would it be to find out whether the dataset is suitable for a project you want to carry out?
    \item Do you identify any aspects of the dataset that you think should be altered?
    \item Any other points you want to make?
\end{itemize}

\begin{thebibliography}{9}

\bibitem{deng2009imagenet} 
Jia Deng, Wei Dong, Richard Socher, Li-Jia Li, Kai Li, and Li Fei-Fei. 
\textit{ImageNet: A large-scale hierarchical image database}. 
In 2009 IEEE Conference on Computer Vision and Pattern Recognition, pages 248--255, 2009.

\bibitem{gebru2021datasheets} 
Timnit Gebru, Jamie Morgenstern, Briana Vecchione, Jennifer Wortman Vaughan, Hanna Wallach, Hal Daum\'{e} III, and Kate Crawford. 
\textit{Datasheets for datasets}. 
Communications of the ACM, 64(12):86--92, 2021. URL: \url{https://arxiv.org/abs/1803.09010}.

\end{thebibliography}

\end{document}
